---
title: "Retail Case Study"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
Customers<-read.csv("Case study 1 (Retail)/Customer.csv")
Prod_Hierarchy<-read.csv("Case study 1 (Retail)/prod_cat_info.csv")
Prod_Hierarchy$Product_Code<-paste(Prod_Hierarchy$prod_cat_code,'-',Prod_Hierarchy$prod_sub_cat_code)
Transactions<-read.csv("Case study 1 (Retail)/Transactions.csv")
Transactions$product_code<-paste(Transactions$prod_cat_code,'-',Transactions$prod_subcat_code)
Customers$customer_Id<-as.character(Customers$customer_Id)
Customers$city_code<-as.character(Customers$city_code)
Customers$DOB<-as.Date(Customers$DOB,format = "%d-%m-%Y")
Transactions$transaction_id<-as.character(Transactions$transaction_id)
Transactions$cust_id<-as.character(Transactions$cust_id)
Transactions$prod_cat_code<-as.character(Transactions$prod_cat_code)
Transactions$prod_subcat_code<-as.character(Transactions$prod_subcat_code)
Transactions$tran_date <- lubridate::dmy(Transactions$tran_date)
Prod_Hierarchy$prod_cat_code<-as.character(Prod_Hierarchy$prod_cat_code)
Prod_Hierarchy$prod_sub_cat_code<-as.character(Prod_Hierarchy$prod_sub_cat_code)

# View(Customers)
# View(Prod_Hierarchy)
# View(Transactions)

# str(Customers)
# str(Transactions)
# str(Prod_Hierarchy)

library(dplyr)
library(lubridate)
library(hablar)
```

#1. Merging  
Using base merge function:
```{r}
m1<-merge(Transactions, Prod_Hierarchy, by.x = "product_code",by.y = "Product_Code", no.dups = FALSE)
#View(m1)

m2<-merge(Customers,m1,by.x="customer_Id", by.y = "cust_id",no.dups = FALSE)
Customers_final<-m2[,c(-8,-9,-15,-17)]
View(Customers_final)
```

Using Dplyr merge:
```{r}
m3<-inner_join(Transactions, Prod_Hierarchy, by= c("product_code" = "Product_Code"))
m4<-inner_join(Customers,m3,by=c("customer_Id" = "cust_id"))
Customer_Final<-m4[,c(-7,-8,-15,-17)]

## Displaying extract of merged data set.
head(Customer_Final)
```

#2. Summary Report  
a.Column names and their data types
```{r}
str(Customer_Final)
```

b.Top/Bottom 10 observations
```{r}
head(Customer_Final, 10)
tail(Customer_Final, 10)
```

c."Five-number summary" for continuous variables
```{r}
Qty<-fivenum(Customer_Final$Qty)
Rate<-fivenum(Customer_Final$Rate)
Tax<-fivenum(Customer_Final$Tax)
Total_Amt<-fivenum(Customer_Final$total_amt)
Five_number_Summary<-data.frame(Qty,Rate,Tax,Total_Amt, row.names=c("Min","1st Quart","Median","3rd Quart","Max"))
print(Five_number_Summary)
```

d.Frequency tables for all categorical variables
```{r}
## Extract of Freq table for customer id
head(table(Customer_Final$customer_Id),50)
## Freq table for Gender
Customer_Final$Gender[Customer_Final$Gender==""]<-NA
table(Customer_Final$Gender)
## Freq table for City
table(Customer_Final$city_code)
## Freq table for Store type
table(Customer_Final$Store_type)
## Freq table for Product Categories
table(Customer_Final$prod_cat)
```

#3.Histogram and Frequency bars
```{r}
hist(Customer_Final$Qty, col="#E4717A", xlab = "Quantity")
hist(Customer_Final$total_amt,border="red", xlab = "Total Amount")
barplot(table(Customer_Final$Gender),xlab = "Gender")
barplot(table(Customer_Final$Store_type),col="light blue",xlab = "Store Type")
```

#4.Transaction Analysis
```{r}
##a.Time period of available transaction data
TimePeriod<-max(Customer_Final$tran_date)-min(Customer_Final$tran_date)
print(TimePeriod)

##b.Count of transactions where total amount was negative
Customer_Final %>% filter(total_amt<0) %>% distinct(transaction_id) %>% count()
```

#5.Category Popularity
```{r}
library(ggplot2)
Customer_Final %>% filter(!is.na(Customer_Final$Gender)) %>% ggplot(aes(prod_cat,fill=Gender))+geom_bar(position='dodge')+labs(x="Product Category")
```

#6.Maximum customers in city
```{r}
d<-Customers %>% group_by(city_code) %>% count()
paste("Maximum customers are from city code",d$city_code[which.max(d$n)])

##Percent of Customers from this city
pct<-max(d$n)/nrow(Customers)
print(pct)
```

#7.Store-type analysis
```{r}
Totals<-Customer_Final %>% filter(Qty>0) %>% group_by(Store_type) %>% summarise(TotalQty=sum(Qty),TotalValue=sum(total_amt))

MaxbyQty<-Totals$Store_type[which.max(Totals$TotalQty)]
paste("Store-type selling max products by Quantity:", MaxbyQty)
MaxbyValue<-Totals$Store_type[which.max(Totals$TotalValue)]
paste("Store-type selling max products by Value:", MaxbyValue)
```

#8.Total amount earned from Electronics and Clothing categories for Flagship stores
```{r}
Customer_Final %>% filter(grepl("Flagship",Store_type) & prod_cat %in% c("Electronics","Clothing")) %>% summarise(Tot_Amt=sum(total_amt))
```

#9.Total amount earned from Males in Electronics category
```{r}
Customer_Final %>% filter(Gender=="M" & prod_cat == "Electronics") %>% summarise(Tot_MnE=sum(total_amt))
```

#10. Customers having more than 10 unique transactions
```{r}
Freq_cust<-Customer_Final %>% filter(total_amt>0) %>% group_by(customer_Id) %>% count()
NoofFreqcust<-nrow(Freq_cust %>% filter(n>10))
print(NoofFreqcust)
```

#11.Customers aged between 25-35  
a. Total amount spent in Electronics and Books
```{r}
Customer_Final$Age<-year(max(Customer_Final$tran_date))-year(Customer_Final$DOB)
Customer_Final %>% filter(between(Age,25,35) & prod_cat %in% c("Electronics","Books")) %>% summarise(AmtSpent=sum(total_amt))
```

b. Total amount spent by above customers between 1st Jan 2014 to 1st Mar 2014
```{r}
Customer_Final %>% filter(between(Age,25,35) & between(tran_date,as.Date("2014-01-01",format = "%Y-%m-%d"),as.Date("2014-03-01",format = "%Y-%m-%d"))) %>% summarise(Amt_Spent=sum(total_amt))
View(Customer_Final)
```

